
\xname{GridIntro}
\begin{Label}{GridIntro}
\introsection{Overview over the Grid Category}
\end{Label} 

{\em 
  NOTE: This documentation follows the pattern of the 
  \xlink{STL Documentation}[\footnote{see {\tt http://www.sgi.com/Technology/STL}}]{http://www.sgi.com/Technology/STL},
  it may be helpful consulting its sections
  \xlink{{\em Introduction\/}}{http://www.sgi.com/Technology/STL/stl_introduction.html}
  and
  \xlink{{\em Usage of the documentation\/}}{http://www.sgi.com/Technology/STL/doc_introduction.html}.
}
      
  Components belonging to the 
  \glossarylink{category}{category}
  {\em Grid\/} provide a uniform 
  interface for geometric grids. 
  As unifying mathematical framework
  we utilize the notion of a
  \xlink{CW-complex}{http://www.treasure-troves.com/math/CW-Complex.html}
  provided by algebraic topology. 
  More common special cases are
  simplicial complexes or triangulations,
  planar straight line graphs (PSLG),
  finite element meshes and Cartesian grids.
  
  We distinguish three fundamental layers of functionality (sub-categories):
  \begin{itemize}
  \item  the \sectionlink{ combinatorial layer }{CombinatorialLayer}
  \item the \sectionlink{ data association layer }{DataAssociationLayer}
  \item the \sectionlink{ geometrical layer }{GeometricalLayer} 
  \end{itemize}
  
  The {\em combinatorial layer\/} views a grid 
  simply as a {\em lattice\/} or, more generally, {\em poset},
  given by the {\em incidence\/} relationship.
  The entities involved are of  a purely discrete nature;
  there is not notion of coordinates of vertices or the like.
  The notion of dimension used here is a combinatorial or {\em intrinsic\/}
  dimension.
  
  The {\em grid function\/} layer adds the possibility to associate data 
  with the discrete grid elements, 
  for example to put marks on vertices,
  or to store solution values on cells.
  
  
  The {\em geometrical layer\/} provides an embedding into a concrete geometrical space of 
  an ({\em external\/}) dimension which is at least the intrinsic dimension. 
  For example, a grid of (intrinsic) dimension 2 could be embedded in 3-space,
  like the surface of a torus.
  

  
  We  informally discuss these layers in the introductory sections
  on the
  \sectionlink{combinatorial}{CombinatorialLayer}, 
 \sectionlink{grid function}{DataAssociationLayer}
 and \sectionlink{geometrical}{GeometricalLayer} layers, respectively.

 A more formal description of the
 \glossarylink{concepts}{concept} involved starts in 
 \sectionlink{the concept section.}{Concepts}
 We will refer to them with a trailing (C).

 In the third part of the documentation, concrete \sectionlink{components}{Components}
 are described, sorted by functionality (algorithms, grid ranges, iterators and so on).
 These components separate into two layers, depending on whether they are `leaf' components
 (for example basic grid implementations),
 or generic extensions, that is, classes which provide additional functionality 
 for each grid component that is a \Glossarylink{model} of a certain \Glossarylink{concept}.
 In fact, most components belong to this second layer,
 which evidently greatly enhances reuse.
  
  \bigsep 

  In the present framework, a mathematical grid is not represented by a {\em single\/}
  entity (e.g. a class), 
  rather, {\em different\/} classes cooperate to this aim.
  This is already visible by the dissection into the three fundamental categories,
  namely combinatorics, data association and geometry.
  
  Moreover, each of these categories introduces many sub-categories 
  and refined concepts, 
  especially the combinatorial one.
  In order to understand how grids are represented by these concepts,
  it is important to understand their mutual relationship.

  For each concrete grid type, there is one class implementing
  the abstract \sectionlink{grid concept}{Grid}.
  {\small (In the STL parlance, it is \glossarylink{{\em models\/}}{model} of the concepts.)}
  To actually use this grid type, one has to access a number of
  different classes accounting for a representation 
  of this mental "The Grid" thing one may have in mind.
  Thus, there are seperate classes for 
  \link{grid elements}{element-intro} (vertices, edges, cells),
  \link{iterators}{sequential-it-intro}, 
  and \sectionlink{grid functions}{DataAssociationLayer}.
  The concrete grid types act as a sort of {\em central entry point\/} 
  giving access to these {\sl associated types\/},
  which one can think of  as  forming a sort of `halo'
  around the central grid type.

  The number of supported types  will not be the same for all grids.
  The basic \conceptlink{Grid}{Grid} concepts requires any types at all
  --- everything useful comes in through refinements of this concept.
  A grid even does not need to define all possible element types ---
  there are grid types used only for I/O 
  (see \sectionlink{\type{IstreamComplex2DFmt}}{istream-complex2d-fmt})
  which do with a quite
  minimalistic set of associated types.
  Most profoundly, grids differ in the amount of
  \link{incidence iteration}[~(see section \Ref)~]{incidence-it-intro}
  they can support --- a trade-off between storage consumption and functionality.


  At first reading, 
  one can consider this type association as fixed.
  But in principle, it is possible to use different `halo' types 
  with one and the same grid type.
  This technique allows a fine control of grid functionality
  which is not attainable with a monolithic approach.

  \label{intro-grid-types}
  An important  technical question is now  
  how to access `halo types' in a consistent and uniform manner. 

  The way chosen here is to collect all non-generic grid-related types into
  a dedicated entity called {\tt grid_types},
   associated to a concrete grid type.
  (With `non-generic', we mean types that have to be specialized for each
  concrete grid type.) 
  In practice, this is just a class template that has to be
  \glossarylink{specialized}{partial specialization} for concrete grids.

  This mechanism provides algorithms with a single entry point to 
  type information (and perhaps other) related to a grid parameter.
  It also offers a clean possibility to pass a different `halo',
  that is, a customized set of types which behave differently.
  For example, one could use special iterators for doing 
  limited algorithm animation in an automatic way.
  Currently, only few algorithms are actually parameterized by 
  grid types, an example is the \type{CGT} parameter of 
  \sectionlink{CalculateNeighborCells}{cell-neighbor-search}.
  
  We may contrast this  with the corresponding mechanisms for STL containers:
  These, too, are associated with some service types, namely the 
  iterator type(s), which are accessed by means of nested typedefs:
  \begin{example}
    Container C;
    typedef Container::iterator iter_type;
    iter_type i = C.begin();
  \end{example}
  In our approach, an analogous action would look like the following:
  \begin{example}
    MyGrid g;
    typedef grid_types<MyGrid> gt; // default set of related types
    gt::VertexIterator vi = g.FirstVertex();
  \end{example}
  A {\em quantitative\/} difference to the STL is, that in the grid category,
  many more types are necessary. 
  The total functionality making up a grid is also more broadly
  distributed among these types.

  An important {\em qualitative\/} difference is the parameterization of algorithms.
  STL algorithms are never parameterized by containers but always by iterators
  forming a logical range {\tt [begin, end)}. This captures well the only aspect of 
  containers the algorithms are interested in, namely that they represent linear 
  sequences.
  In contrast, algorithms for grids are parameterized by grids, and often also
  by grid functions or grid geometries, which correspond  in some sense
  to the STL containers. 
  This is due to the fact that grids in this case act as placeholders
  for the underlying mathematical structures on top of which the algorithm
  performs its tasks.

  The set of types used by these algorithms is rich, corresponding to
  the richer structures of the domain. Moreover, these types are not
  independent, by closely relate to each other.
  Therefore, the responsibility for bundling the right types together
  has to be concentrated somewhere.
  This is in essence what the {\tt grid\_types} mechanism does:
  It maps grid types to a set of related types.
  If one want to control the way  algorithms access grid-related types,
  one has to parameterize them with such mappings, 
  in addition to their regular parameters.
 

  

